---
title: "HW 04"
author: "Megan Hokama"
format: html
toc: true
---

```{r load-packages, message = FALSE}
library(tidyverse)
# add any other packages you use here

if(!require(pacman))
  install.packages("pacman")

pacman::p_load(
  tidyverse,
  dplyr, 
  viridis,
  scales,
  ggplot2,
  RColorBrewer,
  forcats,
  janitor,
  here,
  gganimate,
  gapminder,
  gifski,
  tidytext,
  patchwork,
  colorspace
)

```

## 1 - Degrees over time

```{r}
#| label: loading dataset

ba <- read_csv("data/BA_degrees.csv")

ba |>
  glimpse()


```

Explanation: This side-by-side bar plot allows you to compare the number of Bachelor's degrees awarded in different degree areas in 2015, with each group of bars representing a specific degree area. The proportional difference in lengths between the bars is easy to interpret, providing a clear representation of the relative proportions. Since there are a significant amount of fields, it would be difficult to see the different types on a pie chart in comparsion to the side by side bar plot.

```{r}
#| label: make stacked bar plot
# Filter the dataset for the year 2015
ba_2015 <- subset(ba, year == 2015)

#Create side by side bar plot
ba_2015 <- ba|>
  filter(year == 2015)

ggplot(ba_2015, aes(x = count, y = field))+
  geom_col(fill = "orange")+
  scale_x_continuous(labels = comma)+
  labs(title = 'Bachelor’s Degrees Awarded in 2015', 
       x = 'Degree Amount',
       y = 'Field of Degree')+
  theme(plot.title = element_text(size = 20),
        axis.text = element_text(size = 6))

```

```{r}
#| label: Time series plot

# Filter the dataset for the degree field "Education"
ba_educ <- ba|>
  filter(field == "Education")

# Create a time series plot with a straight line fit
ggplot(ba_educ, aes(x = year, y = perc)) +
  geom_line() +
  geom_smooth(method = "lm", se = FALSE, color = "orange") +  # Add a linear fit
  labs(title = 'Proportion of Education Degrees Over Time', 
       x = 'Year', 
       y = 'Proportion (%)') +
  theme_minimal()


```

Explanation: Based on the time series plot, the proportion of individuals who received an Education degree decreased across time. The straight line fit clearly visualizes the sharp decline from 1970 to 2015.

```{r}
#| label: Top degree plot
# Function to calculate the coefficient of variation
coef_var <- function(x) {
  sd(x)/mean(x)
}

# Calculate the coefficient of variation for each field
variation <- ba |>
  group_by(field) |>
  summarize(coef_var = coef_var(perc))

# Identify the fields that changed the most
top_changed_fields <- variation |>
  top_n(5, wt = coef_var)  # Select top 5 fields with the highest coefficient of variation

# Filter the dataset for the identified fields
highlighted_data <- ba |>
  filter(field %in% top_changed_fields$field)

# Create a color palette for the highlighted fields
color_palette <- brewer.pal(length(top_changed_fields$field), "Set1")

# Create a time series plot with highlighted and gray lines
ggplot() +
  geom_line(data = other_data, aes(x = year, y = perc, group = field), color = "gray", alpha = 1) +
  geom_line(data = highlighted_data, aes(x = year, y = perc, group = field, color = field), size = 1.5) +
  scale_color_manual(values = color_palette) +
  labs(title = 'Proportions of All Degrees Over Time',
       x = 'Year',
       y = 'Proportion (%)') +
  theme_minimal()+
  coord_cartesian(ylim = c(0,.025))

```

Explanation: I defined "changed the most" as the ones with the highest variability or standard deviation in their proportions over the years.

```{r}
#|label: Time series with lumped "other" field

ba_lumped <- ba %>%
  mutate(field = fct_other(as.character(field), keep = top_changed_fields$field))

# Create a time series plot with lumped "Other" field and highlighted fields
ggplot(ba_lumped, aes(x = year, y = perc, group = field, color = field)) +
  geom_line(size = 1) +
  scale_color_manual(values = c(color_palette, "gray"), guide = FALSE) +
  labs(title = 'Proportions of All Degrees Over Time',
       x = 'Year',
       y = 'Proportion (%)') +
  theme_minimal()+
  coord_cartesian(ylim = c(0,.025))
```

Explanation: According to the time series plot, parks & recreation degrees have showed a greater difference over time, while library science degrees showed decrease in change.

2.  **Country populations.**

```{r}
#| label: Race Bar Chart Animation

country <- read.csv(here("data","country-pop.csv"), skip=4)|>
  janitor::clean_names()

country |>
  glimpse()

top10 <- country[, c("country_name", "x2020")]|>
  arrange(desc(x2020)) |>
  head(55)

top10 |>
  glimpse()

top10 <- country|>  
  filter(country_name %in% c("China","India", "United States", "Indonesia",
                             "Pakistan", "Brazil", "Nigeria", "Bangladesh",
                             "Russian Federation", "Mexico"))

top10year <- top10|>
  pivot_longer(cols = 5:65, names_to = "year",
               values_to = "population")|>
  select("country_name", "population", "year")|>
  mutate(year = substring(year, 2))|>
  mutate(year = as.numeric(year))|>
  group_by(year)|>
  arrange(year, -population)|>
  mutate(rank = 1:n())

top10year |>
  glimpse()

staticplot <- ggplot(top10year, aes(rank, group = country_name,
fill = as.factor(country_name), color = as.factor(country_name))) +
  geom_tile(aes(y = population/2,
                height = population,
                width = 0.9), alpha = 0.8, color = NA) +
  geom_text(aes(y = 0, label = paste(country_name, " ")), vjust = 0.2, hjust = 1) +
  geom_text(aes(y = population, label = population, hjust = 0)) +
  coord_flip(clip = "off", expand = FALSE) +
  scale_y_continuous(labels = scales::comma) +
  scale_x_reverse() +
  guides(color = "none", fill = "none") +
  theme(axis.line = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none",
        panel.background = element_blank(),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_line(linewidth = 0.1, color = "grey"),
        panel.grid.minor.x = element_line(linewidth = 0.1, color = "grey"),
        plot.title = element_text(size = 25, hjust = 0.5, face = "bold", colour = "black", vjust = -1),
        plot.subtitle = element_text(size = 15, hjust = 0.5, face = "italic", color = "black"),
        plot.caption = element_text(size = 8, hjust = 0.5, face = "italic", color = "black"),
        plot.background = element_blank(),
        plot.margin = margin(2, 2, 10, 4, "cm"))

anim <- staticplot + transition_states(year, transition_length = 4, state_length = 1) +
  view_follow(fixed_x = TRUE)  +
  labs(title = 'Population per Year with Top 10 Countries : {closest_state}',
       caption  = "Population | Data Source: The World Bank")

animate(anim, 200, fps = 20,  width = 1200, height = 1000,
        renderer = gifski_renderer("gganim.gif"))
```
Explanation: In order to create the racing bar chart, I referenced https://www.r-bloggers.com/2020/01/how-to-create-bar-race-animation-charts-in-r/. This helped me develop a gif that shows the top 10 countries populations over time. China and India indicate a large population increase over time. 

3.  **Battle of the newspapers.**

```{r}
#| label: Newspaper

#Load Data
charl <- read.csv(here("data","charlotte_observer_2019.csv"))

balt <- read.csv(here("data","baltimore_sun_2019.csv"))

#Combine headlines
all_headlines <- bind_rows(
  mutate(charl, newspaper = "Charlotte Observer"),
  mutate(balt, newspaper = "Baltimore Sun")
) %>%
  select(newspaper, title)

# Tokenize the words and remove stop words
all_words <- all_headlines %>%
  unnest_tokens(word, title) %>%
  anti_join(stop_words)

# Get the 20 most common words for each newspaper
common_words_charlotte <- all_words %>%
  filter(newspaper == "Charlotte Observer") %>%
  count(word, sort = TRUE) %>%
  head(20)

common_words_baltimore <- all_words %>%
  filter(newspaper == "Baltimore Sun") %>%
  count(word, sort = TRUE) %>%
  head(20)

# Print the results
print("Common words in Charlotte Observer headlines:")
print(common_words_charlotte)

print("Common words in Baltimore Sun headlines:")
print(common_words_baltimore)

```

Explanation: The words most frequently exhibited by Baltimore or Charlotte indicates common themes found in each region. For Charlotte, common themes include the location and sport's teams such as charlotte, carolina, hornets, panthers. However, Baltimore showed more themes in services and politics such as police, services, school, and trump. One common theme is that Baltimore also had a common theme of naming the location, maryland, baltimore, and md as well. 

```{r}
#| label: Bar chart for common words

# Add the 'newspaper' column back to common_words_charlotte and common_words_baltimore
common_words_charlotte <- common_words_charlotte %>%
  mutate(newspaper = "Charlotte Observer")

common_words_baltimore <- common_words_baltimore %>%
  mutate(newspaper = "Baltimore Sun")

# Create bar charts with rotated x-axis labels
plot_charlotte <- common_words_charlotte %>%
  ggplot(aes(x = fct_reorder(word, n), y = n, fill = factor(newspaper))) +
  geom_col(fill = "orange") +
  labs(title = "Top 20 Common Words in Charlotte Observer Headlines",
       x = "Word",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8),  # Set the size parameter as needed
        plot.title = element_text(size = 8))  # Set the title size as needed

plot_baltimore <- common_words_baltimore %>%
  ggplot(aes(x = fct_reorder(word, n), y = n, fill = factor(newspaper))) +
  geom_col(fill="skyblue") +
  labs(title = "Top 20 Common Words in Baltimore Sun Headlines",
       x = "Word",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8),  # Set the size parameter as needed
        plot.title = element_text(size = 8))  # Set the title size as needed

# Display the plots side by side
plot_charlotte + plot_baltimore + plot_layout(ncol = 2)

```
```{r}
#|label: binding and faceting
# Add a newspaper column to identify the source
charl$newspaper <- "Charlotte Observer"
balt$newspaper <- "Baltimore Sun"

# Identify common variables
common_vars <- intersect(colnames(charl), colnames(balt))

# Select only common variables from each dataset
charl_selected <- charl %>% select(all_of(common_vars))
balt_selected <- balt %>% select(all_of(common_vars))

# Function to find common words
find_common_words <- function(data) {
  data %>%
    select(title) %>%
    unnest_tokens(word, title) %>%
    anti_join(stop_words) %>%
    count(word, sort = TRUE) %>%
    slice_head(n = 10)
}

# Find common words for each newspaper
common_words_charlotte <- find_common_words(charl_selected)
common_words_baltimore <- find_common_words(balt_selected)

# Identify the top 10 words for each newspaper
top10_charlotte <- common_words_charlotte %>% slice_head(n = 10)
top10_baltimore <- common_words_baltimore %>% slice_head(n = 10)

# Combine the top 10 words for each newspaper
top10_combined_data <- bind_rows(
  mutate(top10_charlotte, newspaper = "Charlotte Observer"),
  mutate(top10_baltimore, newspaper = "Baltimore Sun")
)

# Create faceted bar plots with only the top 10 words
facet_bar_plots <- top10_combined_data %>%
  ggplot(aes(x = fct_reorder(word, n), y = n, fill = word)) +
  geom_col() +
  facet_wrap(~ newspaper, scales = "free_y") +
  labs(title = "Top 10 Common Words in Headlines",
       x = "Word",
       y = "Count",
       fill = "Word") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8)  # Adjust size as needed
  ) +
  guides(fill = "none")  # Remove the legend

print(facet_bar_plots)

```
4.  **Brexit.**

```{r brexit data}

brex <- readr::read_csv('data/brexit.csv')

#Filter out "Don't Know" and then adjust region names 
brex %>%
  filter(opinion != "Don't know") %>%
  mutate(
    opinion = factor(opinion, levels = c("Very well", "Fairly well", "Fairly badly", "Very badly")),
    region = case_when(
      region == "scot" ~ "Scotland",
      region == "north" ~ "North",
      region == "midlands_wales" ~ "Midland/Wales",
      region == "rest_of_south" ~ "Rest of South",
      region == "london" ~ "London",
      TRUE ~ as.character(region)
    ),
    region = factor(region, levels = c("London", "Rest of South", "Midland/Wales", "North", "Scotland"))
  ) -> brex

#Plot A
brexa <- brex %>%
  group_by(region, opinion) %>%
  summarise(count = n(), .groups = "drop") %>%
  mutate(pct = count / sum(count) * 100) %>%
  ggplot(aes(x = pct, y = region, fill = opinion)) +
  geom_bar(position = "fill", stat = "identity") +
  guides(fill = guide_legend("Opinion", reverse = TRUE)) +
  scale_x_continuous(labels = scales::percent) +
  scale_fill_discrete_diverging(palette = "Blue-Red") +
  labs(x = "Percent", y = NULL) +
  theme_minimal()


#Plot B
brexb <- ggplot(brex, aes(y = region, fill = opinion)) +
  geom_bar() +
  guides(fill = guide_legend("Opinion", reverse = TRUE)) +
  scale_fill_discrete_diverging(palette = "Blue-Red") +
  labs(x = "Count", y = NULL) +
  theme_minimal()


#Plot C
brexc <- ggplot(brex, aes(y = opinion, fill = opinion)) +
  geom_bar() +
  facet_wrap(~region, ncol = 5) +
  scale_fill_discrete_diverging(palette = "Blue-Red") +
  guides(fill = "none") +
  scale_x_continuous(breaks = seq(0, 200, 100)) +
  labs(
    x = NULL,
    title = "How well or badly do you think the government are doing\nat handling Britain's exit from the European Union?",
    subtitle = "YouGov Survey, 2-3 September 2019",
    caption = "Source: bit.ly/2lCJZVg."
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor.x = element_blank(), 
    axis.title.y = element_blank(), 
    plot.title = element_text(size = 12))
        

(((brexa | brexb) &
   theme(plot.tag.position = c(0, 0.65))) / brexc) +
  plot_annotation(tag_levels = 'A') +
  plot_layout(guides = 'collect') &
  theme(legend.position = 'top')
```
